---
title: "Stat231::Lab4 - Data Viz Part II"
output:
  learnr::tutorial:
  progressive: true
allow_skip: true
runtime: shiny_prerendered
description: >
  For this lab, we will once again be focusing on making clear, consice plots to present data. Instead of using a prebuilt themes, we will be creating our own unique visualizations. 
---
  
```{r, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  collapse = TRUE,
  comment = "#>")

library(learnr)
library(dplyr)
library(magrittr)
library(ggplot2)
library(ggthemes)
library(openintro)
library(inteRactive231)
```
## Custom Data Visualization

For this lab, we will once again be focusing on making clear, consice plots to present data. Instead of using a prebuilt themes, we will be creating our own unique visualizations. 

First, lets start with our plot from lab 3 that included labels.  

```{r labelled-histo, echo = TRUE}
ggplot(data = gpa_study_hours, aes(x = gpa)) +
  geom_histogram(color = "White", binwidth = 0.25) +
  labs( #the labs() function allows us to add labels to our graph
    title = "Distribution of Grade Point Average",
    subtitle = "For 193 Introductory Statistics Students",
    x = "GPA",
    y = "Number of Students",
    caption = "Source: Private US University, 2012"
  )
```

To customize a ggplot image, we'll add a `theme()` function to the end and change specific parts of the graph. Here is a [full list](https://ggplot2.tidyverse.org/reference/theme.html) of what you can change with the `theme()` function. It is a substantial list but don't worry, we won't change all of them.

## Custom Text

First I'm going to adjust my *text*. I'd prefer to have my title a little larger and in Times New Roman font (which R maps to serif). I'd also like the axis labels to be in italics.

```{r custom-text, echo = TRUE}
ggplot(data = gpa_study_hours, 
       aes(x = gpa))+
  geom_histogram(color = "White", binwidth = 0.25)+
  labs(
    title = "Distribution of Grade Point Average",
    subtitle = "For 193 Introductory Statistics Students",
    x = "GPA",
    y = "Number of Students",
    caption = "Source: Private US University, 2012"
  )+
  theme(
    title = element_text(size = 16,
                         family = "serif"),
    axis.title = element_text(face = "italic")
  )
```

## Custom Panel

Next, I'm going to adjust the *panel* the plot is on. I'll change the background color and the grid. 

```{r custom-panel, echo = TRUE}
ggplot(data = gpa_study_hours, 
       aes(x = gpa))+
  geom_histogram(color = "White", binwidth = 0.25)+
  labs(
    title = "Distribution of Grade Point Average",
    subtitle = "For 193 Introductory Statistics Students",
    x = "GPA",
    y = "Number of Students",
    caption = "Source: Private US University, 2012"
  )+
  theme(
    title = element_text(size = 16,
                         family = "serif"),
    axis.title = element_text(face = "italic"),
    panel.background = element_rect(fill = "lightblue"), 
    panel.grid.major.x = element_blank(), 
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(color = "blue"), 
    panel.grid.minor.y = element_blank()
  )
```

## Custom Plot Background

My next step will be to address the *plot* background (the white space behind the blue). I'd like it to be the same color. I'll also adjust the color and fill of the bars. 

```{r custom-plot, echo = TRUE}
ggplot(data = gpa_study_hours, 
       aes(x = gpa))+
  geom_histogram(fill = "aquamarine4",
                 color = "lightblue", 
                 binwidth = 0.25)+
  labs(
    title = "Distribution of Grade Point Average",
    subtitle = "For 193 Introductory Statistics Students",
    x = "GPA",
    y = "Number of Students",
    caption = "Source: Private US University, 2012"
  )+
  theme(
    title = element_text(size = 16,
                         family = "serif"),
    axis.title = element_text(face = "italic"),
    panel.background = element_rect(fill = "lightblue"), 
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(color = "blue"),
    panel.grid.minor.y = element_blank(),
    plot.background = element_rect(fill = "lightblue")
  )
```

## Your Turn!
Create a custom version of your histogram from lab 3. It should include one (or more): 
  
*change to text

*change to panel

*change to plot

*change to histogram/boxplot

```{r custom-data-viz, exercise = TRUE}

```

```{r custom-data-viz-hint}
ggplot(data = gpa_study_hours, aes(x = study_hours)) + 
  geom_histogram(fill = ___,
                 color = ___, 
                 binwidth = ___) +
  labs(
    title = "Study Hours per week", 
    subtitle = "For 193 introductory statistics students",
    x = "Study Hours",
    y = "Number of Students",
    caption = "Source: Private US University, 2012"
  ) +
  theme(
    title = element_text(____),
    panel.background = element_rect(____),
    plot.background = element_rect(____)
  )
```

```{r custom-data-viz-solution}
# NOTE: There are many possible correct solutions. 
ggplot(data = gpa_study_hours, aes(x = study_hours)) + 
  geom_histogram(fill = "goldenrod3",
                 color = "white", 
                 binwidth = 5) +
  labs(
    title = "Study Hours per week", 
    subtitle = "For 193 introductory statistics students",
    x = "Study Hours",
    y = "Number of Students",
    caption = "Source: Private US University, 2012"
  ) +
  theme(
    title = element_text(color = "white"),
    panel.background = element_rect(fill = "red4"),
    plot.background = element_rect(fill = "red4")
  )
```

## Submit

Congratulations! You've completed our fourth lab day! Please enter your name and student ID below, then click `Generate`. Copy/paste the hash code into your submission on Sakai. 

```{r, echo=FALSE, context = "server"}
inteRactive231::encoder_logic()
```

```{r encode, echo=FALSE}
learnrhash::encoder_ui(ui_before = hash_encoder_ui())
```
