---
title: "Stat231::Lab3 - Data Viz Part I"
output:
  learnr::tutorial:
  progressive: true
allow_skip: true
runtime: shiny_prerendered
description: >
  Being able to analyze data is a good quality. Being able to present it in a clear, concise format is even better. In part I of our data viz labs, we create clear and concise plots. 
---
  
```{r, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  collapse = TRUE,
  comment = "#>")

library(learnr)
library(dplyr)
library(magrittr)
library(ggplot2)
library(ggthemes)
library(openintro)
library(inteRactive231)
```
## Quality Data Visualization

Being able to analyze data is a good quality. Being able to present it in a clear, concise format is even better. In part I of our data viz labs, we create clear and concise plots from the `gpa_study_hours` dataset. 

For the examples, we'll use the `gpa` variable. For exercises you will use the `study_hours` variable.

## Base Plot

First, lets draw our base histogram. 

```{r base-histo, echo = TRUE}
ggplot(data = gpa_study_hours, aes(x = gpa)) +
  geom_histogram()
```

Recall, we did make a few adjustments to this in lab 1; we changed the color of the border of the bars. 

```{r border-histo, echo = TRUE}
ggplot(data = gpa_study_hours, aes(x = gpa)) +
  geom_histogram(color = "White", binwidth = 0.25)
```


## Labels

It would definitely be helpful to add a title and clean up the labels a bit. Using the 'Help' tab in the Directory pane, look up our data set, `gpa_study_hours`, and read the description. We'll use this to write our labels. 

```{r labelled-histo, echo = TRUE}
ggplot(data = gpa_study_hours, aes(x = gpa)) +
  geom_histogram(color = "White", binwidth = 0.25) +
  labs( #the labs() function allows us to add labels to our graph
    title = "Distribution of Grade Point Average",
    subtitle = "For 193 Introductory Statistics Students",
    x = "GPA",
    y = "Number of Students",
    caption = "Source: Private US University, 2012"
  )
```

## Style

For our next step, we'll add a little flair. We'll use some styles or "themes" from the `ggthemes` package. Use the 'Packages' tab in the Directory pane to look up `ggthemes` and take a look at some of our options. Note: In part II of the lab we will create more custom styles. You can think of part I (this lab) as using a cookie recipe someone else made. We'll eat the cookie, see if we like it, and change what we don't in the next lab.

```{r theme-histo, echo = TRUE}
ggplot(data = gpa_study_hours, aes(x = gpa)) +
  geom_histogram(color = "White", binwidth = 0.25) +
  labs(
    title = "Distribution of Grade Point Average",
    subtitle = "For 193 Introductory Statistics Students",
    x = "GPA",
    y = "Number of Students",
    caption = "Source: Private US University, 2012"
  ) +
  theme_solarized()
```

I like this theme, but the gray bars seem out of place. I'll change the recipe a bit. 

```{r styled-histo, echo = TRUE}
ggplot(data = gpa_study_hours, aes(x = gpa)) + 
  geom_histogram(fill = "lightgoldenrod2",
                 color = "White", 
                 binwidth = 0.25) +
  labs(
    title = "Distribution of Grade Point Average",
    subtitle = "For 193 Introductory Statistics Students",
    x = "GPA",
    y = "Number of Students",
    caption = "Source: Private US University, 2012"
  ) +
  theme_solarized()
```

That looks much better. To change the color of the bars, I added the `fill` argument to the `geom_histogram` function. Here is a list of [colors](http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf) that you can use. 

##  Your Turn!
Create a histogram for `study_hours` using a different color for the bars and a theme from the ggthemes package.

```{r histogram, exercise = TRUE}

```

```{r histogram-hint}
ggplot(data = gpa_study_hours, aes(x = ___)) + 
  geom_histogram(fill = ___,
                 color = ___, 
                 binwidth = ___) +
  labs(
    title = ___, 
    subtitle = ___,
    x = ___,
    y = ___,
    caption = ___
  ) +
  theme____()
```

```{r histogram-solution}
# NOTE: There are many possible solutions in regards to styling
ggplot(data = gpa_study_hours, aes(x = study_hours)) + 
  geom_histogram(fill = "gold",
                 color = "black", 
                 binwidth = 5) +
  labs(
    title = "Study Hours per week", 
    subtitle = "For 193 introductory statistics students",
    x = "Study Hours",
    y = "Number of Students",
    caption = "Source: Private US University, 2012"
  ) +
  theme_void()
```
Create a boxplot for `study_hours` in a similar fashion. 

```{r boxplot, exercise = TRUE}

```

```{r boxplot-hint}
ggplot(data = gpa_study_hours, aes(x = ___)) + 
  geom_boxplot(fill = ___,
                 color = ___) +
  labs(
    title = ___, 
    subtitle = ___,
    x = ___,
    caption = ___
  ) +
  theme____()
```

```{r boxplot-solution}
# NOTE: There are many possible solutions in regards to styling.
ggplot(data = gpa_study_hours, aes(x = study_hours)) + 
  geom_boxplot(fill = "gold",
                 color = "black") +
  labs(
    title = "Study Hours per week", 
    subtitle = "For 193 introductory statistics students",
    x = "Study Hours",
    caption = "Source: Private US University, 2012"
  ) +
  theme_void()
```

## Submit

Congratulations! You've completed our third lab day! Please enter your name and student ID below, then click `Generate`. Copy/paste the hash code into your submission on Sakai. 

```{r, echo=FALSE, context = "server"}
inteRactive231::encoder_logic()
```

```{r encode, echo=FALSE}
learnrhash::encoder_ui(ui_before = hash_encoder_ui())
```
