---
title: "Stat231::Lab6 - Analysis of Variance Test"
output:
  learnr::tutorial:
  progressive: true
allow_skip: true
runtime: shiny_prerendered
description: >
  For this lab, we will be peaking at section 7.5 (without worry about all the formulas). The analysis of variance - anova - test is equivalent to the chi-squared test but for numeric data instead of categorical data; It will compare the means of three or more samples. 
---
  
```{r, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  collapse = TRUE,
  comment = "#>")

library(learnr)
library(dplyr)
library(magrittr)
library(openintro)
library(ggplot2)
library(inteRactive231)
```

## ANOVA

For this lab, we will be peaking at section 7.5 (without worry about all the formulas). The analysis of variance - anova - test is equivalent to the chi-squared test but for numeric data instead of categorical data; It will compare the means of three or more samples. 

We will be going back to the diamonds data set and running an anova test to see if there is a difference in the average price of diamonds based on clarity. We will nest the test inside a `summary()` function to get the test and print out of results in one line. 


```{r one-line-aov, echo = TRUE}
summary(aov(price ~ clarity, data = diamonds))
```

The important parts of this print out are the $F$ value and $Pr(>F)$. The $F$ value is similar to a $z$ or $t$ score. The $Pr(>F)$ is a p-value. 

Using the p-value method, $<2e-16$ means the p-value is less than $0.0000000000000002$. Using any of the typical values for alpha, we would reject our null hypothesis and can say there is a difference in the average price of diamonds based on their clarity rating.

Using the classic method, we need to calculate a critical F value. 

```{r f-statistic, echo = TRUE}
qf(0.95, #1-alpha
   df1 = 7,
   df2 = 53932)
```

This means our critical region starts at approximately $F = 2$, and our test statistic is $F = 215$. We are well into the critical region and again conclude that there is a difference in the average price of diamonds based on their clarity. 

## Your Turn!
Use the classic *and* p-value method to determine if there is a difference in the average price of diamonds based on the color of the diamond. 

```{r price-aov, exercise = TRUE}

```

```{r price-aov-hint}
# P-Value
summary(aov(___ ~ ___, data = diamonds))

# Conclusion: 

#-------------------------------------------------------------------------------
# Classic
qf(0.95, #1-alpha
   df1 = ____,
   df2 = ____)

# Conclusion
```

```{r price-aov-solution}
# P-Value
summary(aov(price ~ color, data = diamonds))

# Conclusion: Reject the null hypothesis. The average price of a diamond is different depending on color.

#-------------------------------------------------------------------------------
# Classic
qf(0.95, #1-alpha
   df1 = 6,
   df2 = 53932)

# Conclusion: Reject the null hypothesis. The average price of a diamond is different depending on color.
```

## Submit

Congratulations! You've completed our sixth lab day! Please enter your name and student ID below, then click `Generate`. Copy/paste the hash code into your submission on Sakai. 

```{r, echo=FALSE, context = "server"}
inteRactive231::encoder_logic()
```

```{r encode, echo=FALSE}
learnrhash::encoder_ui(ui_before = hash_encoder_ui())
```
